---
title: "Getting started with aquacropr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with aquacropr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

aquacropr drives the FAO AquaCrop **stand-alone plugin** from R. It does not
reimplement the model.

## Workflow

1. Point at the executable (`set_aquacrop_path()`).
2. Describe free scalars in a YAML `parameters:` file (`{CCx}` in a `.CRO`
   template, bounds, `from_file` / `to_file`).
3. `render_templates()` writes concrete AquaCrop text files.
4. `run_aquacrop()` launches the plugin (it reads `LIST/ListProjects.txt`).
5. `read_out()` parses daily `.OUT` files; `objective_spec()` scores them.

A runnable Tunis-wheat script lives next to the package:
`Example_aquacrop/run_wheat.R`.

```{r}
library(aquacropr)
set_aquacrop_path("C:/path/to/AquaCrop73.exe")
set_simul_dir("C:/path/to/plugin/SIMUL")

parameters <- read_parameters("parameters.yaml")
validate_parameters(parameters, template_dir = "templates")
render_templates(parameters, parameter_defaults(parameters),
                 template_dir = "templates", output_dir = "DATA")
run_aquacrop(working_dir = "C:/path/to/plugin")
day <- read_out("C:/path/to/plugin/OUTP/example_wheatPRMday.OUT")
```

## Calibration and sensitivity

`calibrate()` wraps `evaluate_candidate()` in `DEoptim` (default),
`BOBYQA`, `multi-BOBYQA`, `DDS`, or `CMA-ES`. `sa_design()` /
`run_sa_design()` / `plot_sa()` run Morris or Sobol-Jansen. After screening,
pass `names` to `calibrate()` so the rest stay at defaults.

## Designs and metamodels

```{r}
design <- generate_param_design(parameters, n = 30, method = "lhs", seed = 1)
run <- run_param_design(
  design, parameters,
  output_files = c(day = "OUTP/example_wheatPRMday.OUT"),
  working_dir = plugin_dir,
  keep_dir = "metamodel_runs"
)
scored <- score_param_design(run, objective, output_name = "day")
mm <- fit_metamodel(scored, parameters)
validate_metamodel(mm, scored)
calibrate_ego(mm, parameters, objective, sim_file = "OUTP/example_wheatPRMday.OUT",
              nsteps = 10, working_dir = plugin_dir)
```

`create_parameters(c("CCx", "WP"))` scaffolds a starter YAML.
